---
title: "High School Seniors t Test"
output: 
  html_document:
    theme: cerulean
    code_folding: hide
---

```{r, message=FALSE, warning=FALSE}
library(tidyverse)
library(DT)
library(pander)
library(readr)
library(car)
library(mosaic)

HSS <- read_csv("../../Data/HighSchoolSeniors.csv")
#If this code does not work: 
#Use the top menu from RStudio's window to select "Session, Set Working Directory, To Source File Location", and then play this R-chunk into your console to read the HSS data into R. 

## In your Console run View(HSS) to ensure the data has loaded correctly.
```


<!-- Instructions

1. Use the HSS dataset in R to come up with a question that can be answered with a t Test. 

Here is a link to the survey given to the students that participated in the study:
https://ww2.amstat.org/censusatschool/pdfs/C@SQuestionnaire.pdf
(Use the link to learn about what each column of data in the data set represents.)

2. Establish why your question is interesting and rephrase the question using a statistical "null" and "alternative" hypothesis. (Hint, see the "Math 325 Notebook -> R Help -> R-Markdown Hints" page of your textbook for instructions on how to write mathematical hypotheses.)

3. Clearly answer your question using statistics. Be sure to compute a p-value and provide supporting numerical and graphical summaries. If your conclusions don't have numbers in them, they likely are not detailed enough to be meaningful conclusions.

Note: you can create "groups" from the data in many ways. For example, suppose you wanted to create a variable that grouped people according to whether they thought recycling was more important than conserving water. You could do this by:

```{r}
HSS <- HSS %>%
  mutate(recycling_than_water = ifelse(Importance_recycling_rubbish >	Importance_conserving_water, "Recycling More Important", "Water Conservation More Important"))
```

4. Include Q-Q Plots checking the normality of the data and discuss if the requirements of the test are satisfied for your data or not. Note that Q-Q Plots are not graphics, they are diagnostic plots diagnosing the normality of the data.


-->

## Is there a difference between Male and Female Participants in Memory Scores and/or Reaction Times?

### Hypotheses
We understand men and women are different and can preform differently in various areas of achievement. Within our participants in the study is their a difference between how they preformed in memory scores or reaction times?

Null Hypothesis (H0): There is no significant relationship between gender and memory scores and/or reaction times.

Alternative Hypothesis (H1): That there is a difference between male and female participants with memory scores and/or reaction times.

We see from the two data tables we get from favstats that male and female participants scored close to each other. Now we can achieve a better understanding of the data by creating a box plot for each and running both through a t-test to confirm our hypothesis.

```{r}
# View(HSS)

sleep <- (HSS) %>% 
  select(Gender, Ageyears, Score_in_memory_game, Reaction_time, Sleep_Hours_Schoolnight) %>% 
  filter(!is.na(Sleep_Hours_Schoolnight), 
         Sleep_Hours_Schoolnight < 20, 
         !is.na(Score_in_memory_game), 
         !is.na(Reaction_time),
         !is.na(Gender))

sleep_grouped <- sleep %>%
  mutate(Sleep_Hours_Categorical = case_when(
    Sleep_Hours_Schoolnight <=  6 ~ "Group 1 (≤ 6 hours)",
    Sleep_Hours_Schoolnight > 6 ~ "Group 2 (> 6 hours)"
  ))
# View(sleep)


```


### Gender and Memory Score Data
```{r}
favstats(sleep_grouped$Score_in_memory_game ~ sleep_grouped$Gender)

```

### Gender and Reaction Time Data
```{r}
favstats(sleep_grouped$Reaction_time ~ sleep_grouped$Gender)
```


## Gender and Memory Games

```{r}

ggplot(sleep_grouped, aes(x = Gender, y = Score_in_memory_game, fill = Gender)) +
  geom_boxplot() +
  labs(title = "Score in Memory Game by Gender",
       x = "Gender",
       y = "Score in Memory Game") +
  theme_minimal()


t_test_memory <- t.test(Score_in_memory_game ~ Gender, data = sleep_grouped)
print(t_test_memory)



```
### Summary of Results
We see from the data with our p-value of 0.4918 that there is no significant difference in memory game scores between male and female. We are unable to reject the null hypothesis that there is no significant relationship between the number of hours of sleep on school nights and scores in the memory game between gender.



### Gender and Reaction Time

```{r}

ggplot(sleep_grouped, aes(x = Gender, y = Reaction_time, fill = Gender)) +
  geom_boxplot() +
  scale_y_log10() +
  labs(title = "Gender and Reaction Time",
       x = "Gender",
       y = "Reaction Time") +
  theme_minimal()


t_test_reaction <- t.test(Reaction_time ~ Gender, data = sleep_grouped)
print(t_test_reaction)


```
### Summary of Results
By looking at our chart we can see that the two genders are very similar when it comes to reaction time. And as we see in our p-value from the t-test we have a value 0.5676 causing us unable to reject the null hypothesis that there is no significant relationship between male and female and reaction time.

## Conclusion
We are unable to reject both null hypothesis leading to the conclusion that memory scores and reaction times are very similar between men and women in the study. Both p-values from our t-test were greater than 0.05, which means we do not have enough evidence to reject the null hypothesis for either memory scores or reaction times.

