---
title: "Rent"
output: 
  html_document:
    theme: cerulean
    code_folding: hide
    toc: true
    toc_float: true
---

<!-- Welcome to the first Analysis for Math 325! 
The goal of this first assignment is to have you produce a graphical summary and a numerical summary of the "Rent" data set. The fun part is that you get to decide what type of graph and what type of numerical summary to create! You will need to read through the "green text" instructions of this document to find the places where you are supposed to enter R-code in order to complete the assignment. -->


**Week 1 Analysis**

```{r, message=FALSE, warning=FALSE}
# This gray area is called an "R-chunk". 
# These library commands install some powerful functions for your use later on.
library(mosaic)
library(pander)
library(tidyverse)
library(DT) 

# This read_csv command reads in the "Rent" data set into an object called "Rent"
Rent <- read_csv("../Data/Rent.csv")

# To load this data set into your R-Console do the following:
#  1. From your top file menu select "Session -> Set Working Directory -> To Source File Location"
#  2. Press the green "play button" in the top right corner of this gray box (which is called an "R-chunk").
#  3. Then in your "Console" window of 
```

<br />

## Background

Here is a data table showing the available approved housing apartment options at BYU-Idaho for single students. There are 122 entries comprising 57 female and 65 male apartment options.

```{r, warning=FALSE}
# Code to get you started. 
# View(...) works great in the Console, but datatable(...) must be
# used instead within an R-chunk.
datatable(Rent, options=list(lengthMenu = c(3,10,30)), extensions="Responsive")
```


## Graphic

<!-- Present an interesting graphic using the Rent data set. Do not create the same graphics that are shown in the [Good Example Analysis](./Analyses/StudentHousing.html), but you can certainly use those graphics as a guide to inspire you on what to do yourself. -->


```{r}
# Use this R-chunk to...
# ...Create an interesting graphic using the Rent data set.
hist(Rent$Residents, xlab="Number of Residents", main="BYU-Idaho Housing \n One-Semester Entry Level Resident Amounts", col=c("skyblue3","skyblue3","skyblue","skyblue","skyblue","skyblue","skyblue","skyblue"))

```


<!-- Write a few statements here that explain your graphic and what it shows. -->

This graphic shows a basic relationship between the number of residents and the amount of times it appears in the data set. We see that most housing communities are less than 200 per apartment likely due to coding types in Rexburg as larger buildings require stricter coding requirements which costs more to build so building companies opt to build the cheaper option.




```{r}
# Use this R-chunk to...
# ...compute and display a meaningful table of numerical summaries supporting your above graphic.
outputTable <- rbind(Residents=favstats(Rent$Residents))
pander(outputTable[c("min", "Q1", "median", "mean", "Q3", "max", "n")], 
       caption="Resident Amount Summary")


```

<!-- Write a few statements here that introduce your table of numerical summaries and what they show. -->

We see that the minimum amount of students in a housing complex is 6 while our average is around 139. The max is 550 and is likely the larger complexes with either several buildings or the very expansive buildings near downtown Rexburg around Porter Park.


## Conclusion

There are several factors effecting BYU-I student population and as we look at housing we see that companies are opting for smaller more managable size complexes due to a variety of factors from size of property to cost of going above a certain amount of stories.



