---
title: "Stephanie's Housing Request"
output: 
  html_document:
    theme: cerulean
    code_folding: hide
    toc: true
    toc_float: true
execute:
  md: true
---

**Week 2 Analysis**

```{r, message=FALSE, warning=FALSE}
library(mosaic)
library(tidyverse)
library(pander)
library(knitr)
library(DT) # If you get an error stating: 
            # Error in library(DT): there is no package called 'DT'
            # You will need to run: install.packages("DT") 
            # in your Console, then try "Knit HTML" again.
Rent <- read_csv("../Data/Rent.csv")
```

<br />

## Background

------------------------------------------------------------------------

"Hi. My name is Stephanie. I would like to learn about what housing options I have for living at BYU-Idaho next semester. It will be my first semester there, so I would like to find something that is close to campus and around \$300 a month in rent. I'm not too picky on roommates, but I would like somewhere that has a lot of people around so I can get to know as many people as possible. Thanks in advance!"

------------------------------------------------------------------------

## Response

Dear Stephanie,

Below are my top 4 reccomendations for where to live. Based on cost, distance to campus, and number of residents, I have narrowed down the options to the following apartments. 

-------------------------------------------------------------------------

```{r}
# Code to get you started, be sure to use a subset of Rent instead of Rent in this code though.
# datatable(Rent, options=list(lengthMenu = c(3,10,30)), extensions="Responsive")

women_housing <- Rent %>%
  filter(Gender == "F")

women_housing2 <- women_housing %>%
  mutate(
    AvgFloorPlanCostMonthly = AvgFloorPlanCost / 4)

women_housing_main <- women_housing2 %>%
  filter(AvgFloorPlanCostMonthly <= 300)

# View(women_housing_main)

women_housing3 <- women_housing_main %>%
  filter(Residents > (median(women_housing_main$Residents)) &
    CrowFlyMetersToMC <  median(women_housing_main$CrowFlyMetersToMC))

# View(women_housing3)


```

```{r}

women_housing_table <- women_housing3 %>%
  select(Name, AvgFloorPlanCostMonthly, Residents, CrowFlyMetersToMC)
women_housing_table %>%
  kable() #Kable is a way to turn the table into looking cleaner. This way it's more professional to read.

ggplot(women_housing3, aes(x = reorder(Name, -AvgFloorPlanCostMonthly), y = AvgFloorPlanCostMonthly)) +
  geom_col(fill = "steelblue", color = "black") +
  labs(title = "Women's Housing and Monthly Rent",
       x = "Apartment",
       y = "Average Monthly Rent") +
  theme_minimal() +
  coord_flip()


```

-Residents were narrowed down from the given choices to those that had average and above average resident amounts.

```{r}
ggplot(women_housing3, aes(x = reorder(Name, Residents), y = Residents)) +
  geom_col(fill = "steelblue", color = "black") +
  labs(title = "Women's Housing and Residents",
       y = "Total Residents",
       x = "Apartment") +
  theme_minimal() +
  coord_flip()
```


-The green colored dots are the 4 I suggest you look into. The red dots are the ones that are sufficient. The blue dots are the ones that are not as cheap as you would like or farther away then desired. The size of the dots represent the number of residents in the apartment. The table below that is a few other suggestions if for some reason the favorites don't work out for you.


```{r}
# ggplot(women_housing3, aes(x = reorder(Name, -CrowFlyMetersToMC), y = CrowFlyMetersToMC)) +
#   geom_col(fill = "steelblue", color = "black") +
#   labs(title = "Women's Housing and Distance to MC",
#        x = "Apartment",
#        y = "Distance to MC") +
#   theme_minimal() +
#   coord_flip()

womenhousing4 <- women_housing2 %>% 
  filter(Gender == "F")


womenhousing4$chosen <- ifelse(womenhousing4$Name %in% women_housing3$Name, "Favorite",ifelse(womenhousing4$AvgFloorPlanCostMonthly < 300 & womenhousing4$CrowFlyMetersToMC < 618, 
                                      "Desired", 
                                      "Normal"))


meanhousing <- womenhousing4 %>% 
  summarize(mean_cost = mean(AvgFloorPlanCostMonthly), mean_distance = mean(CrowFlyMetersToMC))

belowmean <- womenhousing4 %>% 
  filter(womenhousing4$chosen == "Desired")

belowmean2 <- select(belowmean, c(Name, AvgFloorPlanCostMonthly, CrowFlyMetersToMC, Residents))

# belowmean2 <- round(belowmean2$AvgFloorPlanCostMonthly, 2)
# belowmean2 <- round(belowmean2$CrowFlyMetersToMC, 2)



# meanhousing %>% kable()

ggplot(womenhousing4, aes(x =AvgFloorPlanCostMonthly, y= CrowFlyMetersToMC, color = chosen, size = Residents)) + geom_point()

belowmean2 %>% kable()

```

```{r}
# Creating ranked columns
women_housing3$Rank_AvgFloorPlanCostMonthly <- rank(women_housing3$AvgFloorPlanCostMonthly, ties.method = "min")
women_housing3$Rank_Residents <- rank(women_housing3$Residents, ties.method = "min")
women_housing3$Rank_Distance_MC <- rank(women_housing3$CrowFlyMetersToMC, ties.method = "min")

women_housing3$Average_Row <- rowMeans(women_housing3[, c("Rank_AvgFloorPlanCostMonthly", "Rank_Residents", "Rank_Distance_MC")], na.rm = TRUE)


women_housing3$Average_Row <- round(women_housing3$Average_Row, 2)
```

------------------------------------------------------------------------

## Conclusion

-Below is a ranking for the top apartments suggested with the lower rating being a better housing match for you. On average Riviera is the best option for you but the other 3 are still within your range of what you're looking for. But that is my suggestion for you based on the information requested. Good luck with the apartment search!

-------------------------------------------------------------------------


```{r}
women_ranking_table <- women_housing3 %>%
  select(Name, Rank_AvgFloorPlanCostMonthly, Rank_Residents, Rank_Distance_MC, Average_Row)
women_ranking_table %>%
  kable()

# ggplot(women_housing3, aes(x = reorder(Name, -Average_Row), y = Average_Row)) + #The "-" means to make the order descending. To flip it.
#   geom_col(fill = "steelblue", color = "black") +
#   labs(title = "Average Ranking of Apartments",
#        x = "Apartment",
#        y = "Average Rank") +
#   theme_minimal() +
#   coord_flip()


# View(women_housing3)

```
