---
title: "Food at College"
output: 
  html_document:
    theme: cerulean
    code_folding: hide
---

```{r, warning=FALSE, message=FALSE}
# From your file menu (at the top of R-Studio) select:
# "Session -> Set working directory -> To source file location"
# Then play this chunk to get the data into R.
library(mosaic)
library(car)
library(DT)
library(pander)
library(readr)
library(forcats)
food <- read_csv("../../Data/food.csv") #food.csv is in the Data folder...
```


----

<!-- Instructions:

 Use the Food dataset and a Kruskal-Wallis Test(s) to answer an interesting question(s) that you come up with.
 
 About the dataset:
 
  * GPA - numerical, self-reported current college GPA

  
  * weight - numerical, self-reported weight in lbs


  * gender - "Female", "Male"


  * breakfast - "Cereal", "Donut" 
      The participants were shown a picture of "Cereal" and a "Donut" and asked which one of these pictures they associate with the word "breakfast."


  * calories_chicken - guessing calories in chicken piadina from the options of:
     1 - 265 
     2 - 430 
     3 - 610 
     4 - 720
    (the variable shows the actual number of calories participants selected)


   * calories_day - Importance of consuming calories per day
     1 - i dont know how many calories i should consume 
     2 - it is not at all important 
     3 - it is moderately important 
     4 - it is very important


   * comfort_food - List 3-5 comfort foods that come to mind. 
     Open ended 


   * comfort_food_reasons - What are some of the reasons that make you eat comfort food?      (i.e., anger, sadness, happiness, boredom, etc) - list up to three
     Open ended


   * comfort_food_reasons_coded
     1 - stress
     2 - boredom
     3 - depression/sadness
     4 - hunger
     5 - laziness
     6 - cold weather
     7 - happiness 
     8 - watching tv
     9 - none 


    * cook - how often do you cook?
     1 - Every day 
     2 - A couple of times a week 
     3 - Whenever I can, but that is not very often  
     4 - I only help a little during holidays 
     5 - Never, I really do not know my way around a kitchen


    * cuisine - what type of cuisine did you eat growing up?
     1 - American
     2 - Mexican.Spanish
     3 - Korean/Asian
     4 - Indian
     5 - American inspired international dishes
     6 - other


    * diet_current - describe your current diet
     open ended

	
    * diet_current_coded
     1 - healthy/balanced/moderated/
     2 - unhealthy/cheap/too much/random/
     3 - the same thing over and over
     4 - unclear


    * drink - which picture do you associate with the word “drink”?
     1 - orange juice
     2 - soda 


    * eating_changes  - Describe your eating changes since the moment you got into college?
     Open ended 


    * eating_changes_coded
     1 - worse
     2 - better
     3 - the same
     4 - unclear


    * eating_changes_coded1
     1 - eat faster
     2 - bigger quantity
     3 - worse quality 
     4 - same food
     5 - healthier
     6 - unclear
     7 - drink coffee 
     8 - less food
     9 - more sweets
     10 - timing 
     11 - more carbs or snacking
     12 - drink more water
     13 - more variety


    * eating_out - frequency of eating out in a typical week 
     1 - Never 
     2 - 1-2 times 
     3 - 2-3 times 
     4 - 3-5 times 
     5 - every day


    * employment - do you work? 
     1 - yes full time 
     2 - yes part time 
     3 - no
     4  - other


    * ethnic_food - How likely to eat ethnic food 
     1 - very unlikely 
     2 - unlikely 
     3 - neutral 
     4 - likely 
     5 - very likely 


    * exercise - how often do you exercise in a regular week?
     1 - Everyday 
     2 - Twice or three times per week 
     3 - Once a week
     4 - Sometimes 
     5 - Never


    * father_education - 
     1 - less than high school 
     2 - high school degree 
     3 - some college degree 
     4 - college degree 
     5 - graduate degree 


    * father_profession - what is your father profession?
     Open ended


    * fav_cuisine - What is your favorite cuisine?
     Open ended


    * fav_cuisine_coded
     0-none
     1 - Italian/French/greek
     2 - Spanish/mexican
     3 - Arabic/Turkish
     4 - asian/chineses/thai/nepal
     5 - American
     6 - African 
     7 - Jamaican
     8 - indian


    * fav_food - was your favorite food cooked at home or store bought? 
     1 - cooked at home 
     2 - store bought 
     3 - both bought at store and cooked at home


    * food_childhood - what was your favorite childhood food?
     Open ended



    * fries - which of these pictures you associate with word fries? 
     1 - Mcdonald’s fries
     2 - home fries

 
    * fruit_day - How likely to eat fruit in a regular day 
     1 - very unlikely 
     2 - unlikely 
     3 - neutral 
     4 - likely 
     5 - very likely 


    * grade_level – 
     1 - freshman 
     2 -Sophomore 
     3 - Junior 
     4 - Senior
     
      
    * greek_food - How likely to eat greek food when available?
     1 - very unlikely 
     2 - unlikely 
     3 - neutral 
     4 - likely 
     5 - very likely 
     
      
    * healthy_feel – how likely are you to agree with the following statement: “I feel very healthy!” ?
      1 to 10 where 1 is strongly agree and 10 is strongly disagree - scale
      
      
    * healthy_meal – what is a healthy meal? Describe in 2-3 sentences.
      Open ended
      
      
    * ideal_diet – describe your ideal diet in 2-3 sentences
      Open ended
      
      
    * Ideal_diet_coded
      1 – portion control
      2 – adding veggies/eating healthier food/adding fruit	
      3 – balance
      4 – less sugar
      5 – home cooked/organic
      6 – current diet
      7 – more protein
      8 – unclear
      
      
     * income
      1 - less than $15,000 
      2 - $15,001 to $30,000 
      3 - $30,001 to $50,000 
      4 - $50,001 to $70,000 
      5 - $70,001 to $100,000 
      6 - higher than $100,000
      
     * indian_food – how likely are you to eat indian food when available
      1 - very unlikely 
      2 - unlikely 
      3 - neutral 
      4 - likely 
      5 - very likely 
      
      
     * Italian_food – how likely are you to eat Italian food when available?
      1 - very unlikely 
      2 - unlikely 
      3 - neutral 
      4 - likely 
      5 - very likely 
      
      
     * life_rewarding – how likely are you to agree with the following statement: “I feel life is very rewarding!” ?
      1 to 10 where 1 is strongly agree and 10 is strongly disagree - scale
      
      
     * marital_status
      1 -Single 
      2 - In a relationship 
      3 - Cohabiting 
      4 - Married 
      5 - Divorced 
      6 - Widowed
      
      
     * meals_dinner_friend – What would you serve to a friend for dinner?
      Open ended
      
      
     * mothers_education 
      1 - less than high school 
      2 - high school degree 
      3 - some college degree 
      4 - college degree 
      5 - graduate degree
      
      
     * mothers_profession – what is your mother’s profession? 
      
      
     * nutritional_check - checking nutritional values frequency 
      1 - never 
      2 - on certain products only 
      3 - very rarely 
      4 - on most products 
      5 - on everything
      
      
     * on_off_campus – living situation
      1 - On campus 
      2 - Rent out of campus 
      3 - Live with my parents and commute 
      4 - Own my own house
      
      
     * parents_cook - Approximately how many days a week did your parents cook? 
      1 - Almost everyday 
      2 - 2-3 times a week 
      3 - 1-2 times a week 
      4 - on holidays only 
      5 - never
      
      
     * pay_meal_out - How much would you pay for meal out? 
      1 - up to $5.00 
      2 - $5.01 to $10.00 
      3 - $10.01 to $20.00 
      4 - $20.01 to $30.00 
      5 - $30.01 to $40.00 
      6 - more than $40.01 
      
      
     * Persian_food - How likely to eat persian food when available?
      1 - very unlikely 
      2 - unlikely 
      3 - neutral 
      4 - likely 
      5 - very likely 
      
      
     * self_perception_weight - self perception of weight 
      6 - i dont think myself in these terms 
      5 - overweight 
      4 - slightly overweight 
      3 - just right 
      2 - very fit 
      1 - slim 
      
      
     * Which of the two pictures you associate with the word soup?
      1 – veggie soup
      2 – creamy soup
      
      
     * sports - sports – do you do any sporting activity?
      1 - Yes 
      2 - No 
      99 – no answer
      
      
     * thai_food - How likely to eat thai food when available?
      1 - very unlikely 
      2 - unlikely 
      3 - neutral 
      4 - likely 
      5 - very likely 
      
      
     * tortilla_calories - guessing calories in a burrito sandwhich from Chipolte?
      1 - 580 
      2 - 725 
      3 - 940 
      4 - 1165
      
     * turkey_calories - Can you guess how many calories are in the foods shown below?
     (Panera Bread Roasted Turkey and Avocado BLT)
      1 - 345 
      2 - 500 
      3 - 690 
      4 - 850
      
      
     * type_sports – what type of sports are you involved?
      Open-ended
      
     * veggies_day - How likely to eat veggies in a day? 
      1 - very unlikely 
      2 - unlikely 
      3 - neutral 
      4- likely 
      5 - very likely
      
     * vitamins – do you take any supplements or vitamins?
      1 – yes
      2 – no
      
     * waffle_calories - guessing calories in waffle potato sandwhich 
      1 - 575 
      2 - 760 
      3 - 900 
      4 - 1315


Note this dataset and description come from: © 2020 Kaggle Inc for further details visit:

[Food choices on Kaggle](https://www.kaggle.com/borapajo/food-choices?select=food_coded.csv)


If you want to give your critiquers some ideas about what you have questions on and would like help with, place those details here.-->

<!-- GPA -->

<!-- Eating Changes -->

<!-- healthy_feel – how likely are you to agree with the following statement: “I feel very healthy!” ? -->
<!--       1 to 10 where 1 is strongly agree and 10 is strongly disagree - scale -->

<!-- marital_status -->
<!-- 1 -Single  -->
<!-- 2 - In a relationship  -->
<!-- 3 - Cohabiting  -->
<!-- 4 - Married  -->
<!-- 5 - Divorced  -->
<!-- 6 - Widowed -->

<!-- mothers_education  -->
<!-- 1 - less than high school  -->
<!-- 2 - high school degree  -->
<!-- 3 - some college degree  -->
<!-- 4 - college degree  -->
<!-- 5 - graduate degree       -->

<!-- sports - sports – do you do any sporting activity? -->
<!-- 1 - Yes  -->
<!-- 2 - No  -->
<!-- 99 – no answer -->

<!-- income -->
<!-- 1 - less than $15,000  -->
<!-- 2 - $15,001 to $30,000  -->
<!-- 3 - $30,001 to $50,000  -->
<!-- 4 - $50,001 to $70,000  -->
<!-- 5 - $70,001 to $100,000  -->
<!-- 6 - higher than $100,000 -->

<!-- diet_current_coded -->
<!-- 1 - healthy/balanced/moderated/ -->
<!-- 2 - unhealthy/cheap/too much/random/ -->
<!-- 3 - the same thing over and over -->
<!-- 4 - unclear -->

<!--End your comments to your critiquers before this line.-->
<!--Begin writing your Analysis after this point.-->
----

```{r}
# Let's take a look at the data
# View(food)
```

## Background

### Introduction-Context
In today's day it's becoming harder to afford to eat healthier due to cost. I want to examine if there's a relationship between a higher income and a better diet due to healthier foods becoming more expensive and those of a lesser quality being cheaper and bought more by those with a lower income.
Here's the categories for Income as if you display them on a boxplot the chart become skewed and hard to read:
1 - less than `$`15,000  
2 - `$`15,001 to `$`30,000 
3 - `$`30,001 to `$`50,000 
4 - `$`50,001 to `$`70,000 
5 - `$`70,001 to `$`100,000
6 - higher than `$`100,000


### Hypothesis for ANOVA

$$
H_O: \text{There is no difference in income between the different diet types.}
$$

$$
H_A: \text{There is a difference in income between the different}
$$
$$
a = 0.05
$$


### Numerical/Graphing Summaries


```{r}

#Boxplot and numerical summaries

testdata <- food %>% 
  mutate(diet_current_coded = as.factor(diet_current_coded)) %>% 
  filter(!is.na(income))

testdata <- testdata %>%
  mutate(diet_current_coded = fct_recode(as.factor(diet_current_coded),
                                          "Healthy/Balanced/Moderated" = "1",
                                          "Unhealthy/Cheap/Too much/Random" = "2",
                                          "The same thing over and over" = "3",
                                          "Unclear" = "4"))
# testdata <- testdata %>%
#   mutate(income = fct_recode(as.factor(income),
#                               "Less than $15,000" = "1",
#                               "$15,001 to $30,000" = "2",
#                               "$30,001 to $50,000" = "3",
#                               "$50,001 to $70,000" = "4",
#                               "$70,001 to $100,000" = "5",
#                               "Higher than $100,000" = "6"))

ggplot(testdata, aes(x = diet_current_coded, y = income, color = diet_current_coded)) +
  geom_boxplot() +
  labs(title = "Boxplot of Income by Diet Current Coded",
       x = "Diet Current Coded",
       y = "Income") +
  theme_bw() +
  theme(
    panel.background = element_rect(fill = "lightgrey", color = NA),
    plot.background = element_rect(fill = "lightgrey", color = NA),
    plot.title = element_text(color = "royalblue", face = "bold", size = 16),
    axis.title.x = element_text(color = "royalblue"),
    axis.title.y = element_text(color = "royalblue"),
    legend.background = element_rect(fill = "lightgrey"),
    axis.text.x = element_text(angle = 45, hjust = 1)
  ) +
    coord_flip()



pander(favstats(income ~ diet_current_coded, data = testdata))

```

You can see from the box plot that they do all max out near each other, but the medians are different. The numerical summaries show that the mean income for those with a healthy/balanced/moderated diet is higher than those with an unhealthy/cheap/too much/random diet. The same thing over and over diet is the lowest in income. 

### One-Way Anova


```{r}
# One-Way ANOVA for Sports and diet current coded


aov_sports_diet <- aov(income ~ diet_current_coded, data = testdata)
pander(summary(aov_sports_diet))


```
We have a p-valye of .1471 and due to our alpha of 0.05 we fail to reject the null hypothesis. According to the ANOVA test there is no difference in income between the different diet types.


### qqPlot/Requirements

```{r}
# qqplot to test normality

plot(aov_sports_diet, which = 1:2)


```

We see that the requirements are met by the one-way ANOVA test for the hypothesis. But we're gonna complete a Kruskal-Wallis test to just confirm and compare our assumptions.


### Hypothesis for KW-Test

$$
H_O: \text{There is no difference in income between the different diet types groups.}
$$

$$
H_A: \text{At least one diet grouping has a different income.}
$$

$$
a = 0.05
$$

### Kruskal-Wallis Test

```{r}
#Kruskal-Wallis Test

pander(kruskal.test(income ~ diet_current_coded, data = testdata))

```

The Kruskal-Wallis confirms our rejection of the Null Hypothesis and shows that there is no difference in income between the different diet types.


### Requirements for KW-Test
We can see from prior that the requirements were already met and does not require further testing to check for normality.


### Conclusion
We can conclude from our one-way ANOVA and the Kruskal-Wallis test that we fail to reject the null hypothesis. There is no difference in income between the different diet types. We can see from our p-value that it is greater than 0.05 but it is close and could help us examine the information deeper if we wanted to know more.

















